Comparison of Catalogs

### 2024-07-18 compare/compare.qmd
func compare+scatter

```{python}
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
import glob
import os

dt = QTable(ascii.read("../tables/inner_join.ecsv"), masked=True)
dt.info()
```

```{python}
```

>>>>

```{python}
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
#quantity_support()
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import glob
import os
plt.style.use('ggplot')

dt = QTable(ascii.read("../tables/inner_join.ecsv"), masked=True)
```

The dataset we are going to use consists of `{python} len(dt)` galaxies and `{python} len(dt.colnames)`

# How are we going to compare the data?

## Scatter plots and $R^2$ calculation

1.  [$R^2$]{.underline}: Measures the proportion of variance explained by the linear model.
2.  [Slope of the Fitted Line]{.underline}: Should be close to 1 for a 1-1 correlation.

```{python}
def compare_data(x, y):
    """
    Performs a linear comparison between two datasets.

    This function fits a linear model to the data, calculates the slope and intercept of the fitted line,
    and computes the R-squared value and Pearson correlation coefficient to assess the fit quality.

    Parameters:
    - x (array-like): The independent variable data.
    - y (array-like): The dependent variable data.

    Returns:
    tuple: A tuple containing the following elements:
        - slope (float): The slope of the fitted linear model.
        - intercept (float): The intercept of the fitted linear model.
        - r2 (float): The R-squared value, indicating the proportion of variance explained by the linear model.
        - corr (float): The Pearson correlation coefficient, measuring the linear correlation between x and y.
    """
    x_data = np.ma.array(x.value, mask=np.isnan(x.value))
    y_data =  np.ma.array(y.value, mask=np.isnan(y.value))
    # initialize a linear fitter
    fit = fitting.LinearLSQFitter()

    # initialize a linear model
    line_init = models.Linear1D()

    # fit the data with the fitter
    fitted_line = fit(line_init, x_data, y_data)
    slope = fitted_line.slope.value
    intercept = fitted_line.intercept.value

    # Predict values using the fitted model
    y_pred = fitted_line(x_data)

    # Calculate R-squared
    r2 = r2_score(y_data, y_pred)

    # Calculate Pearson correlation coefficient
    corr, _ = pearsonr(x_data, y_data)

    return slope, intercept, r2, corr

def scatter_plot(x, y, xerr =0, yerr = 0):
    """
    Generates a scatter plot of two datasets with optional error bars, fits a linear model to the data, 
    and displays the fitted line on the plot.

    Parameters:
    - x (array-like with units): Independent variable data.
    - y (array-like with units): Dependent variable data.
    xerr (array-like, optional): Error in the independent variable data. Default is 0.
    yerr (array-like, optional): Error in the dependent variable data. Default is 0.

    Returns:
    None. The function displays a scatter plot with a fitted linear model.
    """
    # Convert data to masked arrays, masking NaN values
    x_data = np.ma.array(x.value, mask=np.isnan(x.value))
    y_data =  np.ma.array(y.value, mask=np.isnan(y.value))

    # Perform linear comparison between the datasets
    c, m, r2, corr = compare_data(x, y)

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    plt.errorbar(x_data, y_data, xerr, yerr, alpha = 0.3, c = "blue", label = "Data", fmt=".k")

    # Plot the fitted line with unit support
    plt.plot(x, c*x + m*x.unit, c = "red", label = "Fit")

    # Set plot labels
    plt.xlabel(f"LCV, {x.info.name} [{x.unit}]")
    plt.ylabel(f"HECATE, {y.info.name} [{y.unit}]")

    # Display legend
    plt.legend()

    # Show the plot
    plt.show()
```

-   Histograms

-   Deviation

# Comparable data

::: panel-tabset
## Coordinates

|  LCV  | HECATE |   Description   | Linear Correlation |
|:-----:|:------:|:---------------:|:------------------:|
| Ra_1  |  RA_2  | Right Ascension |                    |
| Dec_1 | DEC_2  |   Declination   |                    |
|  Dis  |   D    |    Distance     |                    |

## Velocities

|        LCV        |        HECATE         |               Description               | Linear Correlation |
|:----------------:|:----------------:|:-----------------:|:----------------:|
| W50 (with errors) |                       | H I line with at 50% level from maximum |                    |
|       RVel        |    V (with errors)    |      Heliocentric radial velocity       |                    |
|        Vm         |                       |    Amplitude of rotational velocity     |                    |
|        VLG        | V/V_VIR (with errors) |             Radial velocity             |                    |
| cz (with errors)  |                       |          Heliocentric velocity          |                    |

## Morphology and Geometry

|      LCV      |        HECATE        |                        Description                        | Linear Correlation |
|:----------------:|:----------------:|:-----------------:|:----------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |                    |
|      inc      |          I           |                     Inclination (deg)                     |                    |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |                    |

## Luminosities

|   LCV   | HECATE | Description |
|:-------:|:------:|:-----------:|
| logKLum | logL_K |             |
|         |        |             |
|         |        |             |

## Magnitudes

|         LCV         | HECATE | Description |
|:-------------------:|:------:|:-----------:|
| mag_B (with errors) |        |             |
|       mag_FUV       |        |             |
|                     |        |             |

## Absorbsion

|  LCV   | HECATE |                  Description                   |
|:------:|:------:|:----------------------------------------------:|
|   AB   |   AG   |    Galactic extinction/absorption in B band    |
| AB_int |   AI   | Internal/Intrisic B band extinction/absorption |

## SFR

|  LCV   |   HECATE   |                           Description                            |
|:------------------:|:------------------:|:------------------------------:|
|        | logSFR_TIR | Decimal logarithm of the total-infrared SFR estimate \[Msol/yr\] |
|        | logSFR_FIR |  Decimal logarithm of the far-infrared SFR estimate \[Msol/yr\]  |
|        | logSFR_60u |      Decimal logarithm of the 60um SFR estimate \[Msol/yr\]      |
|        | logSFR_12u |      Decimal logarithm of the 12um SFR estimate \[Msol/yr\]      |
|        | logSFR_22u |      Decimal logarithm of the 22um SFR estimate \[Msol/yr\]      |
|        | logSFR_HEC |  Decimal logarithm of the homogenised SFR estimate \[Msol/yr\]   |
|        | logSFR_GSW |       Decimal logarithm of the SFR in GSWLC-2 \[Msol/yr\]        |
| SFRFUV |            |             FUV derived integral star formation rate             |
| SFRHa  |            |          H{alpha} derived integral star formation rate           |

## Masses

|      LCV       |  HECATE  |                        Description                        |
|:------------------:|:------------------:|:------------------------------:|
|     logM26     |          |              Log mass within Holmberg radius              |
|     logMHI     |          |              Log mass within Holmberg radius              |
|                | logM_HEC |      Decimal logarithm of the stellar mass \[Msol\]       |
|                | logM_GSW | Decimal logarithm of the stellar mass in GSWLC-2 \[Msol\] |
| logStellarMass |          |                Stellar Mass from M\*/L=0.6                |
:::
